---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
    author: York Huang
---


## Loading and preprocessing the data

1. Load the data into data frame acts. The data between 8:30 and 9:00 on 2012-11-02 are listed below for example.
```{r}
library(lattice)

acts <- read.csv(unz("activity.zip", "activity.csv"), header=T)
acts[ acts$date == "2012-11-02" & acts$interval >= 830 & acts$interval <= 900, ]

```

2. No extra step is needed to process the data.

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day. Days/Intervals with empty values are ignored.
```{r}
daily <- aggregate( steps~date, data=acts, FUN=sum)
daily
```

2. Make a histogram of the total number of steps taken each day
```{r}
hist(daily$steps, xlab="Steps per day", ylab="Number of days", main="Histograms of total Steps per day")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
daily_mean <- mean(daily$steps)
print(paste("Mean of the total number of steps taken per day =", daily_mean))

daily_median <- median(daily$steps)
print(paste("Median of the total number of steps taken per day =", daily_median))
```

## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
avg_interval <- aggregate( steps~interval, data=acts, FUN=mean)
xyplot( steps~interval, data=avg_interval, type="l", main="Average daily activity pattern")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
max_avg <- avg_interval[ which.max( avg_interval[,2]),]
print(paste("The interval containing maximum steps is" , max_avg$interval, ". And the maximum average steps are", max_avg$steps))
```

## Imputing missing values
There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
num_na = sum(is.na(acts$steps))
print(paste("The total number of rows with NAs is", num_na))
```

2. Use the 5-minute interval mean for the missing interval.


3. Create a new dataset that is equal to the original dataset but with the missing data filled in. The first 6 rows not are filled with data.
```{r}
acts_no_na <- merge(acts, avg_interval, by="interval")

acts_no_na$steps.x <- ifelse(is.na(acts_no_na$steps.x), round(acts_no_na$steps.y), acts_no_na$steps.x)
acts_no_na$steps.y<- NULL
names(acts_no_na) <- list("interval", "steps", "date")

## The following codes just rearrange columns and rows. It is just for getting the same format as the original data.

acts_no_na <- acts_no_na[c("steps", "date", "interval")]
acts_no_na <- acts_no_na[with(acts_no_na, order(date, interval)), ]
head(acts_no_na)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
daily_no_na <- aggregate(steps~date, data=acts_no_na, FUN=sum)
hist(daily$steps, xlab="Steps per day", ylab="Number of days", main="Histograms of total Steps per day (No NA)")

daily_mean_no_na <- mean(daily_no_na$steps)
print(paste("Mean of the total number of steps taken per day (no NA) =", daily_mean_no_na))
print(paste("It is", ifelse(daily_mean== daily_mean_no_na, "the same as", "different to"), "the value with na."))

daily_median_no_na <- median(daily_no_na$steps)
print(paste("Median of the total number of steps taken per day (no NA) =", daily_median_no_na))
print(paste("It is", ifelse(daily_median== daily_median_no_na, "the same as", "different to"), "the value with na."))
```

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
acts_no_na$weekday <- factor((weekdays( as.Date(acts_no_na$date,"%Y-%m-%d")) %in% 
                                 c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')), 
                              levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
head(acts_no_na)
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 

```{r}
avg_interval_no_na <- aggregate(steps~interval+weekday, data = acts_no_na, FUN=mean)
head(avg_interval_no_na)

xyplot( steps~interval|weekday, data=avg_interval_no_na, type="l", layout=c(1,2))
```

